Real-time Attention Assessment for Social Robots

### 2015-08-29 paper.tex
Work on Introduction
A few unresolved issues, marked as 'fix me'

@@

In the recent years, research is pushing forward in building intelligent systems and environments that try to maximize the benefits of cooperation with humans based on verbal and non-verbal communication. The interaction between social robots and humans provide a rich set of behaviors, and reveals a need of capturing the significant passive information provided by the user during the interaction process. Such information should be used to adapt the system to the specific user's needs in order to be able to provide a social or emotional intelligence to the agents.

While new algorithms and robotic technologies ensure better cognitive and interactive abilities, a deeper study about the benefits in the social interaction needs to be done: Is not enough to exhibit human-like capabilities in social tasks, but to do it in a human-like manner. For instance, several studies showed that humans prefer to interact with robots whose task outcome is delayed, shows uncertainty or an undesired output \cite{Admoni,Short}. In fact, the ability to adapt has been proved to be beneficial in long-term interactions \cite{Tielman:2014, Lim:2014}, specifically with children, manifesting a more positive attitude over time.

In order to build intelligent agents, it is necessary to equip them with perceptive capabilities based on sensors an algorithms: face and object tracking and recognition \cite{Zhao:2003, Jafri:2014}, path planning \cite{Galceran:2013}, speech recognition \cite{brick2007incremental} or task learning \cite{calinon2007learning} are some of the current research topics. However, each algorithm is evaluated in their own metric space of its specific problem: efficiency, accuracy or time are some few examples. In \cite{anzalone} is exposed the need to evaluate such processes in terms of their capability to obtain the desired effect in a human-robot interaction context through several proposed metrics around engagement as quality experience indicator.

@@

In this way, in situ Visual Focus of Attention (VFOA) estimation is advantageous to human-robot interaction and a major contribution towards a better understanding of the interaction quality. \cite{yarbus1967eye, barber1976perception} suggest that human's interest is highly correlated with what they look at and the existing relationship between gaze and attention during social interaction has been exposed in \cite{argyle1969social}.

One of the most interesting fields where to develop this research is in a learning context since children's engagement sustainability over time is one of the most important variables to make
the process of learning effective \cite{Umbach}. It is well known that the role of the teacher \cite{smith2005pedagogies}, as well as the way the activity is formulated do matters in the learning efficiency. Moreover, the use of robots in the learning activity presents an opportunity for children
to interact with an embodied, physical agent as part of the learning experience.

Furthermore, while \cite{zhao2012learning} recommends the development of a teachable agent that is actively, rather than passively, engaged in the learning by teaching interaction, the increased learning gain results attained in \cite{okita2006observation} were done so even with the pupil being taught having been instructed to not contribute much information. This suggests that it may not be necessary, from a learning by teaching point of view, for the teachable agent to participate actively in the interaction to incite educational benefits. However, from an engagement point of view, this may not necessarily be true.

In fact, the use of robots in education comes with the potential to engage the child in meta-cognition through the learning by teaching paradigm, wherein a student takes the role of a teacher and experiences stronger educational benefits as a result (such as in \cite{Palinscar1984}).

The focus of this research is to evaluate the quality of the interaction using the VFOA as well as the correctness in the performed task as engagement indicators in the context of the CoWriter project \footnote{The primary repository is \url{https://github.com/chili-epfl/cowriter_letter_learning}.}. In order to do that avoiding the use of 'invasive' technologies such as wearable devices becomes a must to keep the experience as natural as possible.

\section{Related Work}

@@

\section{Research Questions}

>>>>

\usepackage[draft,nomargin,footnote]{fixme}

\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etal}{\textit{et al.}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}

@@

In the recent years, research is pushing forward in building intelligent systems and environments that try to maximize the benefits of cooperation with humans based on verbal and non-verbal communication. The interaction between social robots and humans entails a broad set of situations, and evidences the need for capturing significant backchannel information generated by the user during the interaction process. Such information is used to adapt the system to the specific user's needs to eventually provide a social or emotional intelligence to the agents.

It is indeed not enough to exhibit human-like capabilities in social tasks: a robot needs to achieve human-like manners. For instance, several studies showed that humans prefer to interact with robots whose task outcome is delayed, shows uncertainty or an undesired output \cite{Admoni,Short}. In fact, the ability to adapt has been proved to be beneficial in long-term interactions \cite{Tielman:2014, Lim:2014}, specifically with children, manifesting a more positive attitude over time.

In order to build intelligent agents, it is necessary to equip them with a range of perceptual capabilities: while face and object tracking and recognition~\cite{Zhao:2003, Jafri:2014}, path planning~\cite{Galceran:2013}, speech recognition~\cite{brick2007incremental} or task learning~\cite{calinon2007learning} are amongst the current major research directions, existing literature tends to evaluate each of these algorithms in their own metric space, without considering the overall interaction quality. Along this line, Anzalone \etal~\cite{anzalone} have recently argued in favor of such a global evaluation, and propose to assess these algorithms in term of their capability to obtain the desired effect in a human-robot interaction context, and they correspondingly propose metrics build around the measurement of \emph{engagement} as indicator of the quality of the experience.

@@

In this way, in situ Visual Focus of Attention (VFOA) estimation is advantageous to human-robot interaction and a major contribution towards a better understanding of the interaction quality.\cite{yarbus1967eye, barber1976perception} suggest that human's interest is highly correlated with what they look at and the existing relationship between gaze and attention during social interaction has been exposed in~\cite{argyle1969social}.\fxfatal{This paragraph is not well connected to the previous ones}

\subsection{Importance in learning environments}

Education and learning environments are interesting application fields for research on attention assessment, since sustaining children's engagement over time is a key variable to ensure
effective learning (see for instance~\cite{Umbach}), and they indeed motivate this research. Besides, while~\cite{zhao2012learning} recommends the development of a teachable agent that is actively, rather than passively, engaged in the learning by teaching interaction, the increased learning gain results attained in~\cite{okita2006observation} were done so even with the pupil being taught having been instructed to not contribute much information. This suggests that it may not be necessary, from a learning by teaching point of view, for the teachable agent to participate actively in the interaction to incite educational benefits. However, from an engagement point of view, this may not necessarily be true.\fxfatal{rephrase, and be careful: 'learning by teaching' is only introduced in the next paragraph}

In fact, the use of robots in education comes with the potential to engage the child in meta-cognition through the learning by teaching paradigm, wherein a student takes the role of a teacher and experiences stronger educational benefits as a result (such as in \cite{Palinscar1984}).\fxnote{to be merged with previous paragraph?}

The goal of this research is therefore to \textbf{evaluate the quality of an interaction} using the \textbf{visual focus of attention as an engagement indicator}. We validate our approach with field experiments conducted in the educational context of the CoWriter project~\cite{Hood:2015}, where our technique, being based on a regular, low-end RGB camera, avoids the use of otherwise typically invasive technologies such as eyetrackers, that would negatively impact the natural of the interaction.

\subsection{Related Work}

@@

\subsection{Research Questions}
